---
title: 'Assignment 8.3 : Clustering'
author: "Gourav Verma"
date: "July 27 2019"
output: 
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("C:/Users/f4puslg/Desktop/Garv/ML/Term 2 - R/Assignments")

# Load libraries
library(tidyverse)
library(cluster)
library(factoextra)
library(ggplot2)
library(gridExtra)

# Change scientific expressions
options(scipen = 999)

# Read the datasets
cluster <- read.csv("clustering-data.csv")

```

***
#### **Question a : Plot the dataset using a scatter plot.**
##### **Answer :** Scatter Plot for Clustering dataset.
```{r, echo=FALSE}
ggplot(cluster, aes(x, y)) + 
   geom_point()
```

***
#### **Question b : Fit the dataset using the k-means algorithm from *k=2 to k=12*. Create a scatter plot of the resultant clusters for each value of k.**
##### **Answer :**
```{r, echo=FALSE}

k2 <- kmeans(cluster, centers = 2, nstart = 25)
k3 <- kmeans(cluster, centers = 3, nstart = 25)
k4 <- kmeans(cluster, centers = 4, nstart = 25)
k5 <- kmeans(cluster, centers = 5, nstart = 25)
k6 <- kmeans(cluster, centers = 6, nstart = 25)
k7 <- kmeans(cluster, centers = 7, nstart = 25)
k8 <- kmeans(cluster, centers = 8, nstart = 25)
k9 <- kmeans(cluster, centers = 9, nstart = 25)
k10 <- kmeans(cluster, centers = 10, nstart = 25)
k11 <- kmeans(cluster, centers = 11, nstart = 25)
k12 <- kmeans(cluster, centers = 12, nstart = 25)

p2 <- fviz_cluster(k2, geom = "point", data = cluster) + ggtitle("k = 2")
p3 <- fviz_cluster(k3, geom = "point", data = cluster) + ggtitle("k = 3")
p4 <- fviz_cluster(k4, geom = "point", data = cluster) + ggtitle("k = 4")
p5 <- fviz_cluster(k5, geom = "point", data = cluster) + ggtitle("k = 5")
p6 <- fviz_cluster(k6, geom = "point", data = cluster) + ggtitle("k = 6")
p7 <- fviz_cluster(k7, geom = "point", data = cluster) + ggtitle("k = 7")
p8 <- fviz_cluster(k8, geom = "point", data = cluster) + ggtitle("k = 8")
p9 <- fviz_cluster(k9, geom = "point", data = cluster) + ggtitle("k = 9")
p10 <- fviz_cluster(k10, geom = "point", data = cluster) + ggtitle("k = 10")
p11 <- fviz_cluster(k11, geom = "point", data = cluster) + ggtitle("k = 11")
p12 <- fviz_cluster(k12, geom = "point", data = cluster) + ggtitle("k = 12")

p2
p3
p4
p5
p6
p7
p8
p9
p10
p11
p12

```

***
#### **Question c : Calculate this average distance from the center of each cluster for each value of k and plot it as a line chart where k is the x-axis and the average distance is the y-axis.**
##### **Answer :** 
```{r, echo=FALSE}
fviz_nbclust(cluster, kmeans, k.max = 12, method = "wss")
```

***
#### **Question d : One way of determining the “right” number of clusters is to look at the graph of k versus average distance and finding the “elbow point”. Looking at the graph you generated in the previous example, what is the elbow point for this dataset?**
##### **Answer :** **4** is the elbow point for this dataset.

***
**END**