---
title: 'Assignment 8.2 : Introduction to Machine Learning'
author: "Gourav Verma"
date: "July 27 2019"
output: 
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("C:/Users/f4puslg/Desktop/Garv/ML/Term 2 - R/Assignments")

# Load libraries
library(QuantPsyc)
library(car)
library(caTools)
library(class)
library(caret)
library(e1071)
library(ggplot2)
library(RColorBrewer)

# Change scientific expressions
options(scipen = 999)

# Read the datasets
binary <- read.csv("binary-classifier-data.csv")
binary$label <- factor(binary$label)
trinary <- read.csv("trinary-classifier-data.csv")
trinary$label <- factor(trinary$label)

```

***
#### **Question a : Plot the data from each dataset using a scatter plot.**
##### **Answer :**
```{r, echo=FALSE}

# Set manual color scheme
myColors <- brewer.pal(5, "Set1")
names(myColors) <- levels(trinary$label)
colScale <- scale_colour_manual(name = "grp", values = myColors)

# COde for scatter plot
p1 <- ggplot(binary, aes(x, y)) + 
          geom_point(aes(color = label)) +
          labs(title = "Scatter plot of Binary Classifier Data") + colScale
p2 <- ggplot(trinary, aes(x, y)) + 
          geom_point(aes(color = label)) +
          labs(title = "Scatter plot of Trinary Classifier Data") + colScale
p1
p2
```

***
#### **Question b : Fit a *k* nearest neighbors model for each dataset for *k=3, k=5, k=10, k=15, k=20, and k=25*. Compute the accuracy of the resulting models for each value of k. Plot the results in a graph where the x-axis is the different values of k and the y-axis is the accuracy of the model.**
##### **Answer :** Below are the Accuracies for different values of *k* in Binary dataset. 
```{r, echo=FALSE, warning=FALSE}
set.seed(123)

## Values of k in question
set <- c(3, 5, 10, 15, 20, 25)

## kNN models for Binary dataset

Bsample <- sample(1:nrow(binary), size = nrow(binary) * 0.8, replace = FALSE)
 
Bin.Trn <- binary[Bsample,]
Bin.Tes <- binary[-Bsample,]
 
## Creditability variable dataframe 
Bin.Trn_labels <- binary[Bsample, 1]
Bin.Tes_labels <- binary[-Bsample, 1]

## Apply FOR loop to create KNN and get their accuracy, for all values in 'set' vector.
i = 1
k.Boptm = 1
for (i in set){
  knn.Bmod <- knn(train = Bin.Trn, test = Bin.Tes, cl = Bin.Trn_labels, k = i)
  k.Boptm[i] <- 100 * sum(Bin.Tes_labels == knn.Bmod) / NROW(Bin.Tes_labels)
  k = i
  cat(k, '=', k.Boptm[i], '
  ')
}
```

---

##### Below are the Accuracies for different values of *k* in Trinary dataset. 

```{r, echo=FALSE, warning=FALSE}
## kNN models for Trinary dataset

Tsample <- sample(1:nrow(trinary), size = nrow(trinary) * 0.8, replace = FALSE)

Trin.Trn <- trinary[Tsample,]
Trin.Tes <- trinary[-Tsample,]

## Creditability variable dataframe 
Trin.Trn_labels <- trinary[Tsample, 1]
Trin.Tes_labels <- trinary[-Tsample, 1]

## Apply FOR loop to create KNN and get their accuracy, for all values in 'set' vector.
i = 1
k.Toptm = 1
for (i in set){
  knn.Tmod <- knn(train = Trin.Trn, test = Trin.Tes, cl = Trin.Trn_labels, k = i)
  k.Toptm[i] <- 100 * sum(Trin.Tes_labels == knn.Tmod) / NROW(Trin.Tes_labels)
  k = i
  cat(k, '=', k.Toptm[i], '
  ')
}

```

---
##### Plot for Accuracy matrix for both dataset is shown below.
```{r, echo=FALSE}

set.seed(1)
df <- data.frame(k.Toptm = runif(100, min = 75, max = 100),
                k.Boptm = runif(100, min = 75, max = 100),
                x = rnorm(100))
ylim <- with(df, range(k.Toptm, k.Boptm))

plot(k.Toptm, type = 'b', pch = 20, col = "blue", ylim = ylim, geom = c("line", "point"), xlab = "k - Value", ylab = "Accuracy Level")
points(k.Boptm, type = 'b', pch = 20, col = "red")
legend("bottomleft", c("Binary", "Trinary"),
      col = c("red", "blue"), pch = 19)
```

***
#### **Question c : Looking back at the plots of the data, do you think a linear classifier would work well on these datasets?**
##### **Answer :** I don't think linear classifier will work on these datasets. Decision boundaries between value 0, 1 and 2(in case of Trinary) is not linear. The pots cannot be divided into two pars based on values of label.

***

***
**END**
