---
title: 'Assignment 7.2 : Logistic Regression Model'
author: "Gourav Verma"
date: "July 20 2019"
output: 
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("C:/Users/f4puslg/Desktop/Garv/ML/Term 2 - R/Assignments")
library(QuantPsyc)
library(car)
library(caTools)
library(class)
library(caret)
library(e1071)
options(scipen = 999)
binary <- read.csv("binary-classifier-data.csv")
```
***
#### **Question a : What is the accuracy of the logistic regression classifier?**
```{r, echo=FALSE}
## SPlit the data into training and test
split <- sample.split(binary, SplitRatio = 0.8)
Bin.Tr <- subset(binary, split == "TRUE")
Bin.Te <- subset(binary, split == "FALSE")

## Create Model
Bin.Mod <- glm(label ~ x + y, data = Bin.Tr, family = binomial())

## Create regression
Regs <- predict(Bin.Mod, type = "response")

## Confusion Matrix
confmatrix <- table(Actual_Value = Bin.Tr$label, Predicted_Value = Regs > 0.5)

## Calculate Accuracy
accu <- (confmatrix[(1.1)] + confmatrix[(2.2)]) / sum(confmatrix)
```
##### **Answer :** Below is the summary of the model and **Accuracy** is `r round(accu * 100, digit = 2)`%
```{r, echo=FALSE}
summary(Bin.Mod)
```

---
#### **Question b : How does the accuracy of the logistic regression classifier compare to the nearest neighbors algorithm?**
##### **Answer :** Creating kNN model. Our training dataset has total 1048 rows. Square root of 1048 is 32.37. Hence, creating below 2 kNN models for 'K' of 32 and 33
```{r, echo=FALSE}
## Create Train and Test dataset
set.seed(123)
sample <- sample(1:nrow(binary), size = nrow(binary) * 0.7, replace = FALSE)

Bin.Trn <- binary[sample,]
Bin.Tes <- binary[-sample,]

## Creditability variable dataframe 
Bin.Trn_labels <- binary[sample, 1]
Bin.Tes_labels <- binary[-sample, 1]
```
```{r}
kNN.32 <- knn(train = Bin.Trn, test = Bin.Tes, cl = Bin.Trn_labels, k = 32)
kNN.33 <- knn(train = Bin.Trn, test = Bin.Tes, cl = Bin.Trn_labels, k = 33)
```

+ For kNN.32, below confusion matrix shows Accuracy level of 97%, which is significantly greater than accuracy of logistic regression model.
```{r, echo=FALSE}
confusionMatrix(table(kNN.32, Bin.Tes_labels))
```

+ For kNN.33, below confusion matrix shows Accuracy level of 97%, which is significantly greater than accuracy of logistic regression model.
```{r, echo=FALSE}
confusionMatrix(table(kNN.33, Bin.Tes_labels))
```

***
#### **Question c : Why is the accuracy of the logistic regression classifier different from that of the nearest neighbors?**
##### **Answer :** KNN is a distance based technique while Logistic regression is probability based. We cna consider below points to have difference in accuracy calculated by logistic and kNN.
+ KNN is a non-parametric model, where LR is a parametric model.
+ KNN is comparatively slower than Logistic Regression.
+ KNN supports non-linear solutions where LR supports only linear solutions.
+ LR can derive confidence level (about its prediction), whereas KNN can only output the labels.

***
**END**